---
title: "2022 CRITFC Panel Updates"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: false
---

```{r, message=FALSE, warning=FALSE}
require(tidyverse)
require(magrittr)
require(knitr)
```


We received some updated panel information from CRITFC on February 17th. In this notebook I review the changes from our current panels.

# Chinook Panel

## Add/Subtract Markers

What markers were added or dropped from the three panels in question (Feb 2021 CRITC, December 2021 CRITFC, Ots353 (current SFGL panel))?

### CRITFC Feb 21 -> Dec 21
First let's compare the CRITFC December 2021 panel (update) to the February 2021 panel.
```{r}
critfc_dec21 <- readxl::read_xlsx("Ots Loci Information. BPA. IDT. PROBEseq 20Dec2021.xlsx", sheet = 2)
critfc_feb21 <- readxl::read_xlsx("Ots Loci Information. BPA. IDT. PROBEseq 1Feb2021.xlsx", sheet = 2)

# check which markers ore in the active panel
critfc_dec21 %<>%
  filter(`Last Action` %in% c( "Active Panel", "Testing"))
critfc_feb21 %<>%
  filter(`Last Action` %in% c( "Active Panel", "Testing"))

#markers in dec not in feb
kable(filter(critfc_dec21, !(Assay %in% critfc_feb21$Assay)) %>%
  select(Assay, `Presumed Type`), caption = "New Markers in December Panel")

# markers removed
kable(filter(critfc_feb21, !(Assay %in% critfc_dec21$Assay)) %>%
  select(Assay, `Last Action`), caption = "Removed Markers")
  
```

These don't add up, Feb CRITFC panel lost 4 markers, and gained three, but the total number is the same???  

It's because the marker names under column "assay" aren't consistent. Marker Ots_ZR-575 gets renamed as "Ots_sept9-78" in December. (it looks like this is an error - see primer probe change section)  

So the changes from February to December are as follows

- add 3 AAM markers  
- remove 3 adaptive test markers

### SFGL Current Panel

How does this compare to our current panel? Do we have the Feb2021 Dec2021 or something else?


```{r, message=FALSE}
Ots353 <- read_csv("../Probe_Sequences/Ots353_probeseqs.csv", col_names = FALSE)

Ots353 %<>%
  rename(marker=X1)

kable(Ots353 %>%
  filter(!(marker %in% critfc_dec21$Assay)) %>%
  select(marker), caption = "Markers in Ots353 not in Dec21 CRITFC")


kable(critfc_dec21 %>%
  filter(!(Assay %in% Ots353$marker)) %>%
  select(Assay), caption = "Markers in Dec21 CRITFC, not in Ots353")

```

__CRITFC missing__  
Our panel has 9 markers that CRITFC doesnt have (note ZR-575 IS in the CRITFC panel under a different name). These are the three test markers they removed in their latest round, plus our SWFSC RoSA markers (stat wtih Ots37124)

__SFGL missing__  
Our panel is missing 7 markers that are in the newest CRITFC panel (note sex marker in table above because of the way I pulled these down).   
4 are markers we excluded during our optimization because they were read hogs or had primer-dimer issues. Interestingly 3 of these are flagged in the CRITFC spreadsheets with "keep at Jon's request"  
3 are the new AAM markers.

## Primer/Probe/Correction_value Changes

Let's see if any primer and probe sequences are different.
```{r}
pp_compare <- Ots353 %>%
  rename(`A1-Probe` = X4, `A2-Probe` = X5, `FWD Primer` = X6, Correction = X7, Correction2 = X8) %>%
  select(marker,`A1-Probe`, `A2-Probe`, `FWD Primer`, Correction, Correction2 ) %>%
  left_join(select(critfc_dec21, Assay,`A1-Probe`, `A2-Probe`, `FWD Primer`, Correction, Correction2), by = c("marker" = "Assay"), suffix = c("_SFGL", "_critfc_dec"))

# find differences
pp_compare %<>%
  mutate(change_probe = case_when(`A1-Probe_SFGL` !=  `A1-Probe_critfc_dec` ~ TRUE,
                             `A2-Probe_SFGL` !=  `A2-Probe_critfc_dec` ~ TRUE,
                             TRUE ~ FALSE)) %>%
  mutate(change_correction = case_when(Correction_SFGL !=  Correction_critfc_dec ~ TRUE,
                             Correction2_SFGL !=  Correction2_critfc_dec ~ TRUE,
                             TRUE ~ FALSE)) %>%
  mutate(change_primer = case_when(`FWD Primer_SFGL` !=  `FWD Primer_critfc_dec` ~ TRUE,
                                   TRUE ~ FALSE))

# table of differences
kable(pp_compare %>%
        filter(change_correction == TRUE | change_primer == TRUE | change_probe == TRUE) %>%
      select( sort(current_vars())) %>%
        relocate(marker, starts_with("change"))) 
```

Three markers have different allele correction values and one has a primer and probe change between our panel and the current CRITFC panel.


The primer change for Ots_sept9-78 is completely different and doesn't even align to our primer or it's reverse complement. This is because Ots_sept9-78 appears twice in the critfc excel spreadsheet, and one of the rows is the pesky Ots_ZR-575 marker. Perhaps this is an error in their spreadsheet?

Yes, everything else in the offending row corresponds to Ots_ZR-575. It looks like this is an error in the spreadsheet.

